---
title: "Income Analysis based on college major"
author: "Rohit Benny Abraham"
date: "7/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The goal is to check for a significant association, if any, between college 
major category and the income level. There are 16 different categories and 173 
observations.

## Loading the college dataframe

```{r loading}
library(collegeIncome)
data(college)
head(college)
unique(college$major_category)
```

## Variable transformation

```{r preprocessing}
college$major <- as.factor(college$major)
college$major_code <- as.factor(college$major_code)
college$major_category <- as.factor(college$major_category)
```

## Relationship visualization

```{r EDA}
boxplot(median/1000 ~ major_category, data = college, main = "Income vs. Major",
        ylab="Income (in $1000)", xlab = "Major Category", las = 2)
```

The distribution is not normal based on the boxplot but for practicing linear
model let's assume it to be normal

## Analysis

```{r Analysis}
# First we will reorder the major categories alphabeticaly
college <- college[order(college$major_category), ]

# Applying the linear model with Income as outcome and major as predictor
mdl <- lm(median ~ major_category - 1, data = college)
summary(mdl)$coef

# Apparently, most majors have similar income, except the Business(the highest)
major_category_bus <- relevel(college$major_category, "Business")
mdlBus <- lm(median ~ major_category_bus, data = college)
summary(mdlBus)$coef

business_diff <- summary(mdlBus)$coef[-1,]

# Check for the lowest(significant) p-values
business_diff[order(business_diff[,4])[1:6], ]

# Income for business is significantly different from Computers & Mathematics, 
# Education, Engineering, Humanities & Liberal Arts and Social Science

# We will do the same for the lowest income category i.e. Interdisciplinary
major_category_inter <- relevel(college$major_category, "Interdisciplinary")
mdlInter <- lm(median ~ major_category_inter, data = college)
summary(mdlInter)$coef

inter_diff <- summary(mdlInter)$coef[-1,]

# Check for the lowest(significant) p-values
inter_diff[order(inter_diff[,4])[1:5], ]

# Income for Interdisciplinary is insignificant to other categories
```

The highest income group was of Business with a median salary of $49153.85. 
So by setting income for Business as a major category I ran the linear model and
found that it is significantly different from Computers & Mathematics, 
Education, Engineering, Humanities & Liberal Arts, and Social Science by 
verifying the p-values and keeping the error rate as 5%.

## Gender comparision

Now we will be adding Gender(per_men and per_women) as predictor along with 
major category to check for any significant assocaiation.

```{r Gender}
# Gender comparision with business as the major category
mdlGen <- lm(median ~ major_category + perc_men + perc_women, data = college)
summary(mdlGen)$coef
Gen_diff <- summary(mdlGen)$coef[-1,]
# Check for the lowest(significant) p-values
Gen_diff[order(Gen_diff[,4])[1:5], ]
```

No significant association can be seen due to high p-values

With Business being the major category we will perform the liner modeling again

```{r GenBus}
mdlBusGen <- lm(median ~ major_category_bus + perc_men + perc_women, 
                data = college)
summary(mdlBusGen)$coef

businessGen_diff <- summary(mdlBusGen)$coef[-1,]
# Check for the lowest(significant) p-values
businessGen_diff[order(businessGen_diff[,4])[1:5], ]
```

Again, no significant association can be seen due to high p-values so we can 
conclude that gender has no significant role in income levels for a given major
category